DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-08-05 2015CompleteDerivation/completeDerivation.tex
Second pass at kinematcs

Reductionism by itself would not help us describing a system if the state and the evolution of the pieces were more complicated to study. In this section we further focus our attention to those system that not only are infinitesimally reducible, but their internal dynamics can be ignored. That is, we will assume that the motion in space is necessary and sufficient to reconstruct the state and its dynamics. What we'll find is that the dynamics of each infinitesimal part follows classical Lagrangian mechanics. Furthermore, the Hamiltonian of the system is constrained to the one of a free particles under scalar and vector potential forces.

@@

The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of laws of evolution, is crucial in physics. In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link. In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).

Here we make the link conceptually explicit by assuming the following:

\begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
	For the given system, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. state trajectories in state space).
\end{assump}

@@

	The assumption clearly cannot be taken in general since many systems do not posses this trait. Yet, as this assumption allows to recover the standard link between kinematic and dynamical quantities, there is something fundamental about it. That is: if we assume that the system has no discernible internal dynamics, then the only dynamics that needs to be described is the one associated with its trajectory in space: kinematic equivalence holds.

	This is in fact consistent with observation, though elementary particles are quantum fields therefore some caveats are in order. Massless particles should be excluded as they cannot be localized, which makes it hard to talk about a proper trajectory. Spin is not associated to kinematic quantities: nothing is thought to be actually spinning. As such we can regard it as the internal dynamic of the system. A massive spin $0$ system is one that can be thought as localized and as having no internal structure. For such system we would expect kinematic equivalence to hold and it does: from kinematic momentum we can reconstruct conjugate momentum and vice-versa.

	Therefore, with the appropriate caveat that we are talking about a quantum particle which we haven't yet introduced, the simplest systems are the ones that just describe a trajectory in space. This raises the opposite question: why all systems have \emph{at least} to describe a trajectory in space? This seems motivated by the previous assumption of reducibility: spatial extent is the one way that all materials are reduced. Therefore all particles of all materials, regardless of whether they are elementary or not, are will include spatial degrees of freedom.

	It should be also noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations. This means that the remaining internal dynamics must be identified scalar, vector or tensor quantities. Spin is, in some sense, a quantized version of those.

	While we can make an argument why kinematic equivalence is important because it is valid for elementary system, we cannot actually make an argument why nature should be reducible to elementary system. While it fits our physical intuition, we have no a priori argument for it. We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states. For example, most macroscopic object are neutral though the parts are charged. We could imagine that, while dividing into parts, we find more and more types of charges and forces. We do not have an argument why this cannot happen.
\end{rationale}

@@

As we need to describe trajectories, we introduce $\mathcal{M}$ as the manifold that defines space and time. We call $x^i$ the spatial coordinates, $t$ the time coordinate and $x^\alpha = (t, x^i)$ their combination. A trajectory will be given by a set of functions $x^\alpha(s)$. What we need to understand is the link between these trajectories and the complete state space of the particles of a classical material.

We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$. Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen. Under the kinematic assumption, these units must also be sufficient to define the dynamics. Therefore $\mathcal{M}$ has conceptually two functions: it is the space where the motion unfolds and it is the space that defines the system of units for our states. This already tells us that the our state space is a hyper-surface of $T^*\mathcal{M}$.

@@

Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $(t, q^i) = x^\alpha$. This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable. The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories. The magnitude of the four velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical. Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories. The combined position and velocity, which we call \emph{initial conditions}, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$. This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$. We cannot add more curves, as a homeomorphism preserve dimensionality, therefore we found the suitable space for our trajectories.

Mathematically, under the kinematic assumption the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.

\begin{prop}\label{initial_conditions}
	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifolds over which trajectories are defined. The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
\end{prop}

@@

Being able to convert the state variable for the particles into initial conditions: we also need to be able to express the distributions of the composite states over position and velocity. To do that, we need to be able to count trajectories in the same way we are able to count states. There must exist a metric $g$, consistent with $\omega$, that allows to calculate the cardinality of possibilities for initial conditions. That is: $\omega(dt,dq^i,dE,dp_i) \propto g(dx^\alpha,du^\alpha)$.

Since the both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$. This means that a each direction in space-time $dx^\alpha$ is associated with a direction in the space of velocities $du^\alpha$ to form an independent d.o.f. Across independent d.o.f. we have $\omega(dq^i, dp_j) = 0 =\omega(dq^j, dp_i)$. Therefore if $dx^i$ and $dx^j$ do not form a d.o.f., $g(dx^i, du^j) = 0 = g(dx^j, du^i)$: $g$ is reflexive. We also know that $g$ cannot be alternating, that is $g(dx^i, du^i) \neq 0$: a choice of units for $x^i$ is a choice of units for $u^i$ and therefore they cannot belong to different d.o.f. This means $g$ is symmetric. $\mathcal{M}$ is a pseudo-Riemannian manifold.

Since $g$ is symmetric, it can be diagonalized: $\omega= dq^i \wedge dp_i - dt \wedge dE \propto g = dx^i g_{ii} du^i + dt g_{00} du^0$. This tells us $g_{ii} > 0$ and $g_{00} < 0$. We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities. Since time must use a different unit, we set $g_{00}=-c^2$. Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.

@@

The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of relativity. Yet, they have acquired a different meaning. $g$ does not measure space-time distances: it measure cardinality of possibilities, like $\omega$. Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f.. Space and time, instead, are not independent and the geometry between them is different. In other words: $g$ inherits it's properties from $\omega$.

@@

This also allows to understand conceptually why the speed of a deterministic and reversible signal cannot be greater than $c$. Consider the trajectory of a point particle. Over an interval $dt$ it will travel $dx$ in space. Since the motion is deterministic and reversible, it will travel through all the states defined in the interval $dx$. It will also travel through every moment in time, which corresponds to $cdt$ possibilities. The number of states that will travel in space cannot exceed the number of states it travels in time as it can only be at one position for each instant. Therefore we have:
\begin{align*}
\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
\frac{|dx|}{|dt|} & \leq c \\
\end{align*}
Note, though, that this limit applies to deterministic and reversible motion. It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$. Non-local quantum effects, such as the ones features in EPR experiments, are not examples of deterministic and reversible motion. Therefore they do not have anything to violate: the maximum speed, simply put, does not apply.

@@

The link between $\omega$ and $g$ allows us to constrain the kind of homemorphism between state variables and initial conditions. If we express the possibilities in terms of both state variables and initial conditions we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$. Since  $p_\alpha=p_\alpha(x, u)$, we have $dp_\alpha = m du_\alpha$, where $m$ is the constant of proportionality. We recognize $m$ as the inertial mass. Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary function. This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.

This constrains the form of the extended Hamiltonian. $\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$. Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha(x))g^{\alpha\beta}(p_\beta-A_\beta(x))+V(x)$ where $V(x)$ is an arbitrary function. We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization) for which the modulus of the velocity is constant. We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$. $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha(x)g^{\alpha\beta}(p_\beta-A_\beta(x) + m^2 c^2)$. Applying the same procedure in the time independent case yields the Hamiltonian $H=\frac{1}{2m}(p^i-A^i(x))(p_i-A_i(x))+V(x)$.

@@

The above discussion gives the inertial mass a new physical meaning: it is the conversion constant that tells how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities. In other words: the more massive the particles, the more states per unit velocity. Intuitively, a more massive body is harder to accelerate because it has to go through more states for the same change in velocity. Note that the Hamiltonian excludes massless particles and this makes absolute sense. A massless particle always has the same velocity, $c$, therefore the trajectory is not enough to tell us the dynamics. The motion of the photon is not enough to tell us its energy or its momentum. Therefore it makes sense that, by assuming kinematic equivalence we implicitly required massive particles.

We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve the same purpose: convert one quantity to another preserving the number of possibilities. $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum. $dx = c dt$ converts between ranges of space and time. $dp = m du$ converts between ranges of velocity and conjugate momentum. They allow to keep proper book keeping when mapping sets of states. There is a difference, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material. That is, different homogeneous materials can have different inertial mass.

Another important thing to notice is that the relationship between conjugate momentum and velocity, $p_\alpha = m u_\alpha + A_\alpha(x)$, includes an arbitrary function. As it is arbitrary, we cannot give $p$ a well defined physical meaning in terms of the space-time trajectory. Suppose, in fact, we changed $A$ and $p$ while leaving $u$ unchanged, the physical description in terms of trajectories remains the same. Therefore we should ask whether any choice of $A$ actually represent a different case leading to different trajectories.

The simplest and more insightful way to study this problem is simply to use $x^\alpha$ and $u_\alpha$ as state variable. They are not going to be conjugates of each other but they do capture the entire physical description. Any change in $A$ that does not affect $x^\alpha$ and $mu_\alpha$ does not change the physics. Expressing the Hamiltonian and the Poisson brackets in terms of the initial conditions we have:
\begin{align*}
\mathcal{H}&=\frac{1}{2m}(mu_\alpha mu^\alpha + m^2 c^2) \\
\{x^\alpha, x^\beta\} &= 0 \\
\{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
\{mu_\alpha, mu_\beta\} &= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
\end{align*}
Note how the Poisson bracket between kinetic momentum is the force tensor. The weaker the force, the more $(x^i, mu^i)$ become independent d.o.f. The force tensor, then, tells how much $(x^i, mu^i)$ fail to be independent d.o.f.. With these expressions, the commutator between any function of position and kinetic momentum can now be calculated. For example:
\begin{align*}
d_s x^\alpha &= \{x^\alpha, \mathcal{H}\} = u^\alpha \\
d_s (mu^\alpha) &= \{mu^\alpha, \mathcal{H}\} = F^{\alpha\beta} u_\beta
\end{align*}
where we recognize the covariant form of the Lorentz force.

Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the motion of any function of position and velocity. Such transformation is what we call a gauge transformation.

As we mentioned before, as the extended Hamiltonian is quadratic in the energy, negative energy states are possible. As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$. The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard or anti-particle state. Expanding the equation for the force, we have:
\begin{align*}
\pm d_\tau (\pm m d_\tau x^\alpha) &= m d_\tau^2 x^\alpha = \pm  F^\alpha_\beta d_\tau x^\beta
\end{align*}
Therefore the force act on anti-particle states in the opposite way: they act as they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.

@@

To sum up, the kinematic assumption has more inherent consequences one would expect at first. It means the particles are identified by position and velocity, the system of equation is second order. It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$. Such link allows state space metric $\omega$ to induce the space-time metric $g$. It allows us to define Lagrangians. It constrains the motion to massive particles under scalar and vector potential forces. Moreover, it allowed to re-understand basic concepts, like invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f. independence.

We now have all the building block of classical mechanics, the consequences of three seemingly simple assumption. But that's not all: we have placed ourselves in a very good position to move forward to quantum systems as well by using the new physical insights we have gained.

>>>>

We will now focus our attention to the simplest possible classical material: one for which the motion in space is necessary and sufficient to reconstruct the state and its evolution. That is, the internal dynamics can be ignored. What we'll find is that the dynamics of each infinitesimal part follows classical Lagrangian mechanics. Furthermore, the Hamiltonian of the system is constrained to the one of a free particles under scalar and vector potential forces.

@@

The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of state evolution and its causes, is crucial in physics. In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link. In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).

Here we make the link conceptually explicit with the following assumption:

\begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
	For the given classical material, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. states and their evolution).
\end{assump}

@@

	The assumption clearly cannot be taken in general. Not all material will be elementary in the sense that they may have some internal dynamics. That is: position in space is not their only state variable. Yet, the assumption is fundamental because what we learn about the link between kinematic and dynamical quantities applies to all materials, even the non elementary ones.

	The first natural question to ask is: why do all systems have \emph{at least} to describe a trajectory in space? This seems motivated by the previous assumption of reducibility: spatial extent is the one way that all materials are reduced. Therefore the description of all particles of all materials, regardless of whether they are elementary or not, will include spatial degrees of freedom. Kinematic variables will form a set of independent degrees of freedom that are universal.

	When the system is not elementary, the particles will be described by the kinematic variables plus some additional independent state variables. Yet, the set of all possible evolutions will include those for which kinematic variables evolve independently from the other. For example, the evolutions for which the internal d.o.f. do not change. In other words: the overall dynamics will be the spatial degrees of freedom coupled with the internal degrees of freedom. Therefore by studying an elementary material, we are also studying a component of the dynamics that is present in all materials.

	It should be also noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations. This means that the remaining internal dynamics must be identified by scalar, vector or tensor quantities. This may be used in the future to justify the extension of this work to field theories.

	To be clear: we are not saying that kinematic equivalence is fundamental because everything must reduce to a material with no internal dynamics. While this statement seems reasonable and fits our physical intuition, we have no a priori argument for it. We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states. For example, most macroscopic object are neutral though the parts are charged. We could imagine that, while dividing into parts, we find more and more types of charges, forces and internal dynamics. We do not have an argument why this cannot happen.

	What we are saying is that any classical material is \emph{at least} an elementary material and will have posses the dynamic of one coupled with the internal dynamics.
\end{rationale}

@@

As we need to describe trajectories, we introduce $\mathcal{M}$ as the $n+1$ dimensional manifold that defines space and time. A point in $\mathcal{M}$ will be identified by $x^\alpha = (t, x^i)$, its time and space coordinates. A trajectory will be given by a set of functions $x^\alpha(s)$ where $s$ is an arbitrary parametrization of the curve. What we need to understand is the link between the state space of all the possible trajectories and the complete state space of the particles of a classical material.

We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$. Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen. Under the kinematic assumption, these units must also be sufficient to describe the dynamics. Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states. This already tells us that the our state space is a hyper-surface of $T^*\mathcal{M}$.

@@

Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $q^\alpha = (t, q^i) = x^\alpha$. This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable. The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories. The magnitude of the four velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical. Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories. The combined position and velocity, which we call \emph{initial conditions}, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$. This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$. We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we found the suitable space for the trajectories of particles of a classical material under the kinematic assumption.

Being able to identify particle through both state variables and initial conditions is not enough: we also need to be able to express the distributions of the composite states over position and velocity. As such, the differentiable distribution over state variables must become a differentiable distribution over initial condition. This tells us the transformation is differentiable.

Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.

\begin{prop}\label{initial_conditions}
	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifold over which trajectories are defined. The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
\end{prop}

@@

Not only we must be able to express distributions in terms of both state variable and initial conditions, but integration has to carry over as well. To do that, we need to be able to count trajectories in the same way we are able to count states. There must exist a metric $g$, consistent with $\omega$, that allows to calculate the cardinality of possibilities for initial conditions. That is: $\omega(dt,dq^i,dE,dp_i) \propto g(dx^\alpha,du^\alpha)$.

Since the both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$. To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms an independent d.o.f.. Across independent d.o.f. we have $\omega(dq^i, dp_j) = 0 =\omega(dq^j, dp_i)$. Therefore if $dx^i$ and $du^j$ do not form a d.o.f., $g(dx^i, du^j) = 0 = g(dx^j, du^i)$: $g$ is reflexive. We also know that $g$ cannot be alternating, that is $g(dx^i, du^i) \neq 0$: a choice of units for $x^i$ is a choice of units for $u^i$ and therefore they cannot belong to different d.o.f. This means $g$ is symmetric. $\mathcal{M}$ is a pseudo-Riemannian manifold.

Since $g$ is symmetric, it can be diagonalized: $\omega= dq^i \wedge dp_i - dt \wedge dE \propto g = dx^i g_{ii} du^i + dt g_{00} du^0$. This tells us $g_{ii} > 0$ and $g_{00} < 0$. We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities. Since time must use a different unit, we set $g_{00}=-c^2$ where $c$ is the constant that allows to change an range in time to a range in space of equal cardinality of possibilities (i.e. that counts the same number of states). Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.

@@

The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of relativity. Yet, they have acquired a different meaning. $g$ does not measure space-time distances: it measure cardinality of possibilities, like $\omega$. Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f.. Space and time, instead, are not independent and the geometry between them is different. In other words: $g$ inherits it's signature from $\omega$.

@@

This also allows to understand conceptually why the speed of a deterministic and reversible signal cannot be greater than $c$. Consider the trajectory of a point particle. Over an interval $dt$ it will travel $dx$ in space. Since the motion is deterministic and reversible, it will travel through all the states defined in the interval $dx$. It will also travel through every moment in time, which corresponds to $cdt$ possibilities. The number of states that will travel in space cannot exceed the number of states it travels in time as it can only be at one position for each instant. Therefore we have:
\begin{align*}
\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
\frac{|dx|}{|dt|} & \leq c \\
\end{align*}
Note, though, that this constraint applies only to deterministic and reversible motion. It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$. Non-local quantum effects, such as the ones features in EPR experiments, are not examples of deterministic and reversible motion. Therefore they do not have anything to violate: the maximum speed does not apply to them.

@@

The link between $\omega$ and $g$ allows us to constrain the kind of diffeomorphism between state variables and initial conditions. If we express the possibilities in terms of both state variables and initial conditions we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$. Since  $p_\alpha=p_\alpha(x, u)$, we have $dp_\alpha = m du_\alpha$, where $m$ is the constant of proportionality. We recognize $m$ as the inertial mass. Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary function. This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.

This constrains the form of the extended Hamiltonian. $\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$. Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V(x)$ is an arbitrary function. We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization) for which the modulus of the velocity is constant. We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$. $\mathcal{H}=\frac{1}{2m}((p^\alpha-A^\alpha)(p_\alpha-A_\alpha) + m^2 c^2)$. Applying the same procedure in the time independent case yields the Hamiltonian $H=\frac{1}{2m}(p^i-A^i)(p_i-A_i)+V$.

@@

The above discussion gives the inertial mass a new physical meaning: it is the conversion constant that tells us how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities. In other words: the more massive the particles, the more states per unit velocity. Intuitively, a more massive body is harder to accelerate because it has to go through more states for the same change in velocity. Note that the Hamiltonian excludes massless particles because the kinematic assumption does that implicitly. A massless particle always has the same velocity, $c$, therefore the trajectory is not enough to tell us the dynamics. The motion of the photon is not enough to tell us its energy or its momentum. Therefore it makes sense that, by assuming kinematic equivalence we implicitly required massive particles.

We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve the same purpose: convert one quantity to another preserving the number of possibilities. $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum. $dx = c dt$ converts between ranges of time and space. $dp = m du$ converts between ranges of velocity and conjugate momentum. They allow to keep proper book keeping when mapping sets of states. Using natural units means measuring all these quantities through the cardinality of possibilities. There is a difference, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material. That is, different homogeneous materials can have different inertial mass.

Another important thing to notice is that the relationship between conjugate momentum and velocity, $p_\alpha = m u_\alpha + A_\alpha(x)$, includes an arbitrary function. As it is arbitrary, we cannot give $p$ a well defined physical meaning in terms of the space-time trajectory. Suppose, in fact, we changed $A$ and $p$ while leaving $u$ unchanged, the physical description in terms of trajectories remains the same.

We should also ask whether any choice of $A$ actually represent a different case leading to different trajectories. The simplest and more insightful way to study this problem is to use $x^\alpha$ and $mu_\alpha$ as state variable. They are not going to be conjugates of each other but they do capture the entire physical description. Any change in $A$ that does not affect the relationship between $x^\alpha$ and $mu_\alpha$ does not change the physics. Expressing the Hamiltonian and the Poisson brackets in terms of the initial conditions we have:
\begin{align*}
\mathcal{H}&=\frac{1}{2m}(mu_\alpha mu^\alpha + m^2 c^2) \\
\{x^\alpha, x^\beta\} &= 0 \\
\{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
\{mu_\alpha, mu_\beta\} &= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
\end{align*}
Note how the Poisson bracket between kinetic momentum is the force tensor. The weaker the force, the more $(x^i, mu_i)$ become independent d.o.f. The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f.. With these expressions, the commutator between any function of position and kinetic momentum can now be calculated. For example:
\begin{align*}
d_s x^\alpha &= \{x^\alpha, \mathcal{H}\} = u^\alpha \\
d_s (mu^\alpha) &= \{mu^\alpha, \mathcal{H}\} = F^{\alpha\beta} u_\beta
\end{align*}
where we recognize the covariant form of the Lorentz force.

Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the motion of any function of position and velocity. Such transformation is what we call a gauge transformation. Note the same group of transformations is found by fixing $A$ and changing $p$ such that the laws of motion don't change. In other words: $A$ and $p$ are defined relatively to the each other. $A$ defines the $0$ for $p$ and vice versa. The gauge group expresses the arbitrariness of such definition.

The extended Hamiltonian we found is quadratic in energy, therefore negative energy states are possible. These correspond to the particle anti-states mentioned in \ref{antistates}. As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$. The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard particle state or an anti-state. Expanding the equation for the force, we have:
\begin{align*}
\pm d_\tau (\pm m d_\tau x^\alpha) &= m d_\tau^2 x^\alpha = \pm  F^\alpha_\beta d_\tau x^\beta
\end{align*}
Therefore the force act on anti-particle states in the opposite way: they act as if they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.

@@

To sum up, the kinematic assumption has more inherent consequences one would expect at first. It means the particles are identified by position and velocity and therefore the system of equation is second order. It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$. Such link allows state space metric $\omega$ to induce the space-time metric $g$. It allows us to define Lagrangians. It constrains the motion to massive particles under scalar and vector potential forces. Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f. independence.

We now have all the building block of classical mechanics, the consequences of three seemingly simple assumption. But that's not all: we have also placed ourselves in a very good position to move forward to quantum systems by using the new physical insights we have gained.
